---
title: "municipal city ~ foreclosure"
date: "October 3, 2016"
output: 
  html_document: 
    css: bodycss.css
---

### Data Preparation
* Code used to generate the dataset is available at https://github.com/dimuthu999/muni/blob/master/data%20merge.R
* First list of all issues were identified from MSRB (wrds) table. 326,000 issues after removing tranches which have the same security description and the dated date.
* Detailed description of each of the 326,000 issues were manually obtained from emma.org using the cusip of each issue.
* The description of each issue was used to filter out issues such as school districts, public parks etc.
* issue level information for the remaining issues were obtained from bloomberg. Same information were applied to the securities excluded in the original 326,000 using the security description and the dated date.
* Foreclosure percentage and default percentages were calculated for each zipcode-month using the freddie mac's loanmonth dataset
* Trade data for issues where bloomberg's INDUSTRY_SUBGROUP=="Municipal-City" were downloaded from wrds. default data, foreclosure data and issue level data were matched with the trade data to create the dataset.

```{r warning=FALSE,message=FALSE}
rm(list=ls())
library(plyr)
library(dplyr)
library(stargazer)
setwd("E:/Muni")
muni_city <- readRDS("city_muni_data_merged.rds")
muni_city['yield_minus_tbill'] <- muni_city$YIELD-muni_city$tbill_3month
muni_city['purpose'] <- ifelse(muni_city$MUNI_PURPOSE %in% c("REFUNDING BONDS","REFUNDING NOTES"),"refunding",
                               ifelse(muni_city$MUNI_PURPOSE %in% c("PUBLIC IMPS."),"public improvment","other"))
muni_city['issue_type'] <- ifelse(muni_city$MUNI_ISSUE_TYP %in% c("GENERAL OBLIGATION LTD")
                                  ,"general obligation limited",
                               ifelse(muni_city$MUNI_ISSUE_TYP %in% c("GENERAL OBLIGATION UNLTD"),
                                      "general obligation unlimited","other"))
muni_city['offering_type'] <- ifelse(muni_city$MUNI_OFFERING_TYP %in% c("COMPETITIVE"),"competitive",
                               ifelse(muni_city$MUNI_OFFERING_TYP %in% c("NEGOTIATED"), "negotiated","other"))
muni_city$MUNI_ISSUE_SIZE <- as.numeric(muni_city$MUNI_ISSUE_SIZE)
muni_city['TRADE_YEAR']<-substr(muni_city$TRADE_DATE,1,4)

issues <- muni_city[!duplicated(muni_city[,c("CUSIP.x")]),]
glimpse(muni_city)

```

### Descriptive Stats
```{r warning=FALSE}
#Number of issues
length(unique(muni_city$CUSIP.x))
#Number of trades
nrow(muni_city)
#Summary Price, yield, yield minus tibill rate
stargazer(
  muni_city[, c("DOLLAR_PRICE","YIELD","yield_minus_tbill","forc_pct","default_pct")], type = "text", 
  summary.stat = c("mean", "sd","min", "p25", "median", "p75", "max")
)

#Insurance
table(issues$INSURANCE_STATUS)
#Callable
table(issues$CALLABLE)
#Credit Enhancements
table(issues$MUNI_STATE_TAX)
#Purpose
table(issues$purpose)
#Issue Type
table(issues$issue_type)
#Offering Type
table(issues$offering_type)
#Issue Size
summary(issues$MUNI_ISSUE_SIZE)
```

### Basic Regressions
Marginal Effects: 
* 1 standard deviation increase in foreclosure rate (by 0.001), increases the yeild premium by 11.5% at mean
* 1 standard deviation increase in default rate (by 0.01),increases the yeild premium by 6.5% at mean

```{r results='asis'}
ols <- list()
ols[[1]]<-lm(yield_minus_tbill~forc_pct+factor(purpose)+factor(issue_type)+factor(offering_type)+log(MUNI_ISSUE_SIZE)
             +factor(INSURANCE_STATUS)+factor(CALLABLE)+factor(STATE_CODE)+factor(TRADE_YEAR),data = muni_city)
ols[[2]]<-lm(yield_minus_tbill~default_pct+factor(purpose)+factor(issue_type)+factor(offering_type)+
               log(MUNI_ISSUE_SIZE)+factor(INSURANCE_STATUS)+factor(CALLABLE)+factor(STATE_CODE)+factor(TRADE_YEAR),data = muni_city)
stargazer(ols,column.labels=c('Foreclosure','Default'), omit = c("STATE_CODE","TRADE_YEAR"),omit.labels = c("State","Trade Year"),
          type = "html",no.space=TRUE)

```

